\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-08-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_arXiv.tex
adding arxiv version

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least \emph{statistically} comparable (in the order of magnitude sense) to the physical dynamics timescale. In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time. The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in latency, power and memory footprint of the hardware platform.}}

@@

{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, with high probability, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time. For the Case I in Fig \ref{fig:CompTimeIEEE14bus}, few instances when computational times are $O(10^{-2})$ do occur but are statistically rare.}}

@@

Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig. \ref{fig:WassJointPDFSynthetic} shows the 2-Wasserstein metric between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$. In other words, Fig. \ref{fig:WassJointPDFSynthetic} can be viewed as a nonparametric absolute error curve between $\rho^{\rm{MC}}$ and $\varrho^{k}$. Each snapshot value in Fig. \ref{fig:WassJointPDFSynthetic} is computed by solving the well-known optimal transport linear program (see e.g., \cite[Sec. 4.2-4.3]{halder2014probabilistic}, \cite[Ch. 3]{peyre2019computational}) using the corresponding time-varying weighted point clouds. {\blue{The dynamics helps guide the concentration of the time-varying point clouds in regions of high and low likelihoods in the state space, and hence as time progresses, the empirical Monte Carlo approximants get closer to the proximal updates. Consequently, Figs. \ref{fig:RelErrSynthetic}, \ref{fig:RelErrCovSynthetic} and \ref{fig:WassJointPDFSynthetic} show that after initial transients, the distances between the Monte Carlo and the proximal statistics decrease with time.}}

>>>>

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that the computational timescale is at least \emph{statistically} comparable (in the order of magnitude sense) to the physical dynamics timescale. In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time. The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in latency, power and memory footprint of the hardware platform.

@@

In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, with high probability, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time. For the Case I in Fig \ref{fig:CompTimeIEEE14bus}, few instances when computational times are $O(10^{-2})$ do occur but are statistically rare.

@@

Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig. \ref{fig:WassJointPDFSynthetic} shows the 2-Wasserstein metric between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$. In other words, Fig. \ref{fig:WassJointPDFSynthetic} can be viewed as a nonparametric absolute error curve between $\rho^{\rm{MC}}$ and $\varrho^{k}$. Each snapshot value in Fig. \ref{fig:WassJointPDFSynthetic} is computed by solving the well-known optimal transport linear program (see e.g., \cite[Sec. 4.2-4.3]{halder2014probabilistic}, \cite[Ch. 3]{peyre2019computational}) using the corresponding time-varying weighted point clouds. The dynamics helps guide the concentration of the time-varying point clouds in regions of high and low likelihoods in the state space, and hence as time progresses, the empirical Monte Carlo approximants get closer to the proximal updates. Consequently, Figs. \ref{fig:RelErrSynthetic}, \ref{fig:RelErrCovSynthetic} and \ref{fig:WassJointPDFSynthetic} show that after initial transients, the distances between the Monte Carlo and the proximal statistics decrease with time.
